---
title: "Survival analysis"
author: "Huaibin Ge"
date: "5/23/2019"
output: html_document
---


## 1. Load data
Dataset obtained from https://www.nature.com/articles/nature21386. 
```{r}
library(readxl)
library(tidyverse)
data <- read_excel("Supplemental Table 1-CESC_Data_Summary_Table_Supplement_V2.xlsx", sheet = "supplemental table")
```

## clean data
I rename those variables as the original one except for titile like "CLIN" and strange symbol like ``. Then I assign renamed variables to their desired type of numeric or factor. 
```{r}
data_surv = data %>% 
  rename(batch_number = `CLIN:batch_number`,
         age_at_initial_pathologic_diagnosis = `CLIN:age_at_initial_pathologic_diagnosis`,
         vital_status = `CLIN:vital_status`,
         days_to_last_followup = `CLIN:days_to_last_followup`,
         days_to_last_known_alive = `CLIN:days_to_last_known_alive`,
         days_to_death = `CLIN:days_to_death`,
         Hys_LN = `CLIN:Hys_LN`,
         Hys_LVSI_LN = `CLIN:Hys_LVSI_LN`,
         race = `CLIN:race`,
         ethnicity = `CLIN:ethnicity`,
         Dx_merged = `CLIN:Dx_merged`,
         HPV_status = `CLIN:HPV_status`,
         HPV_Hcall = `CLIN:HPV_Hcall`,
         HPV_Hclade = `CLIN:HPV_Hclade`,
         HPV_clade_4 = `CLIN:HPV_clade_4`,
         HPV_int1 = `CLIN:HPV_int1`,
         clinStage = `CLIN:clinStage`,
         dx_method = `CLIN:dx_method`,
         LNposTF = `CLIN:LNposTF`,
         lymphovascular_invasion_indicator = `CLIN:lymphovascular_invasion_indicator`,
         tumorStatus2 = `CLIN:tumorStatus2`,
         UCEC_like = `CLIN:UCEC-like`,
         Purity_Absolute = `SAMP:Purity_Absolute`,
         Ploidy_Absolute = `SAMP:Ploidy_Absolute`,
         EMT_score = `SAMP:EMT_score`,
         e6_cat_k4 = `CLIN:e6_cat_k4`,
         e6ratio_cat_k4 = `CLIN:e6ratio_cat_k4`,
         e6spl_cat_k4 = `CLIN:e6spl_cat_k4`,
         e6sum_cat_k4 = `CLIN:e6sum_cat_k4`,
         E6_counts_combined = `CLIN:E6_counts_combined`,
         E6_spliced_normalized_counts =`CLIN:E6_spliced_normalized_counts`,
         E6_unspliced_normalized_counts = `CLIN:E6_unspliced_normalized_counts`,
         Ratio_E6_unspliced_to_spliced = `CLIN:Ratio_E6_unspliced_to_spliced`,
         MutSig_rateNon_perMb = `SAMP:MutSig_rateNon_perMb`,
         MutSig_rateSil_perMb = `SAMP:MutSig_rateSil_perMb`,
         MutSig_rateTot_perMb = `SAMP:MutSig_rateTot_perMb`,
         hyperMut = `SAMP:hyperMut`,
         APOBEC_Category = `SAMP:APOBEC_Category`,
         APOBEC_MutLoad_MinEstimate = `SAMP:APOBEC_MutLoad_MinEstimate`,
         somatic_mutations_called = `SAMP:somatic_mutations_called`,
         mRNAseq_data = `SAMP:mRNAseq_data`,
         miRNAseq_data = `SAMP:miRNAseq_data`,
         RPPA_data = `SAMP:RPPA_data`,
         Meth_data = `SAMP:Meth_data`,
         CN_data = `SAMP:CN_data`,
         mRNAseq_k3 = `SAMP:mRNAseq_k3`,
         miRNA_k6 = `SAMP:miRNA_k6`,
         RPPA_k3 = `SAMP:RPPA_k3`,
         CIMP_call = `SAMP:CIMP_call`,
         CN_cluster = `SAMP:CN_cluster`,
         iCluster_All_k3 = `SAMP:iCluster_All_k3`,
         iCluster_Squam_k2 = `SAMP:iCluster_Squam_k2`,
         iCluster_Adeno_k2 = `SAMP:iCluster_Adeno_k2`,
         PARADIGM_k4 = `SAMP:PARADIGM_k4`,
         APOBEC1_339 = `GEXP:APOBEC1:339`,
         APOBEC3A_200315 = `GEXP:APOBEC3A:200315`,
         APOBEC3B_9582 = `GEXP:APOBEC3B:9582`,
         APOBEC3C_27350 = `GEXP:APOBEC3C:27350`,
         APOBEC3D_140564 = `GEXP:APOBEC3D:140564`,
         APOBEC3F_200316 = `GEXP:APOBEC3F:200316`,
         APOBEC3G_60489 = `GEXP:APOBEC3G:60489`,
         APOBEC3H_164668 = `GEXP:APOBEC3H:164668`
         
         ) %>% 
  mutate(vital_status = as.factor(vital_status),
         days_to_last_followup = as.numeric(days_to_last_followup),
         days_to_death = as.numeric(days_to_death),
         Hys_LN = as.factor(Hys_LN),
         Hys_LVSI_LN = as.factor(Hys_LVSI_LN),
         race = as.factor(race),
         ethnicity = as.factor(ethnicity),
         Dx_merged = as.factor(Dx_merged),
         HPV_status = as.factor(HPV_status),
         HPV_status = as.factor(HPV_status),
         HPV_Hcall = as.factor(HPV_Hcall),
         HPV_Hclade  = as.factor(HPV_Hclade),
         HPV_clade_4 = as.factor(HPV_clade_4),
         HPV_int1 = as.factor(HPV_int1),
         clinStage = as.factor(clinStage),
         dx_method = as.factor(dx_method),
         LNposTF  = as.factor(LNposTF),
         lymphovascular_invasion_indicator = as.factor(lymphovascular_invasion_indicator),
         tumorStatus2 = as.factor(tumorStatus2),
         UCEC_like = as.factor(UCEC_like),
         e6_cat_k4 = as.factor(e6_cat_k4),
         e6ratio_cat_k4 = as.factor(e6ratio_cat_k4),
         e6spl_cat_k4 = as.factor(e6spl_cat_k4),
         e6sum_cat_k4 = as.factor(e6sum_cat_k4),
         E6_counts_combined = as.numeric(E6_counts_combined),
         E6_spliced_normalized_counts =as.numeric(E6_spliced_normalized_counts),
         E6_unspliced_normalized_counts = as.numeric(E6_unspliced_normalized_counts),
         Ratio_E6_unspliced_to_spliced = as.numeric(Ratio_E6_unspliced_to_spliced),
         hyperMut = as.factor(hyperMut),
         APOBEC_Category = as.factor(APOBEC_Category),
         somatic_mutations_called = as.factor(somatic_mutations_called),
         mRNAseq_data = as.factor(mRNAseq_data),
         miRNAseq_data = as.factor(miRNAseq_data),
         RPPA_data = as.factor(RPPA_data),
         Meth_data = as.factor(Meth_data),
         CN_data = as.factor(CN_data),
         mRNAseq_k3 = as.factor(mRNAseq_k3),
         miRNA_k6 = as.factor(miRNA_k6),
         RPPA_k3 = as.factor(RPPA_k3),
         CIMP_call = as.factor(CIMP_call),
         CN_cluster = as.factor(CN_cluster),
         iCluster_All_k3 = as.factor(iCluster_All_k3),
         iCluster_Squam_k2 = as.factor(iCluster_Squam_k2),
         iCluster_Adeno_k2 = as.factor(iCluster_Adeno_k2),
         PARADIGM_k4 = as.factor(PARADIGM_k4)
         )

#summary(data_surv)
```

Preprocess
Select curious variable and transform categorical variable into dummy variable becasue Lasso can only deal with continuos variable or dummy variable. 
```{r}
library(caret)
library(glmnet)
library(corrplot)
library(survminer)
library(survival)
# Capture covariates
Cov <- data.frame(data_surv[,-c(1:2, 5, 7, 9, 22, 27:34, 41:49, 51:ncol(data_surv))])

# Deal with factors
dmy = dummyVars("~.", data = Cov)
x_dummy = data.frame(predict(dmy, newdata = Cov))
x_dummy = na.omit(x_dummy)
x_dummy$age_at_initial_pathologic_diagnosis = cut(x_dummy$age_at_initial_pathologic_diagnosis, breaks = c(20, 39, 59, 79), right = FALSE)
x_dummy1 = x_dummy[which(x_dummy$days_to_last_known_alive>0),]
x_dummy1 = na.omit(x_dummy1)
y_input = Surv(x_dummy1$days_to_last_known_alive, x_dummy1$vital_status.Dead)
x_dummy1 = x_dummy1 %>% select(-c(vital_status.Alive, vital_status.Dead, days_to_last_known_alive, age_at_initial_pathologic_diagnosis))
x_input <- model.matrix(~., data = x_dummy1)[,-1]
```


### 2. LASSO to select variables
```{r}
#fit = glmnet(x_input, y_input, family = "cox", alpha = 1)
cv.lasso = cv.glmnet(x_input, y_input, type.measure = "mse", family = "cox")
plot(cv.lasso)

# Select best model
fit = glmnet(x_input, y_input, family = "cox", lambda = cv.lasso$lambda.min)

# Manual selection
fit$beta[fit$beta[,1]!=0,]
selected = c("Hys_LN.LNneg","clinStage.II",
             "lymphovascular_invasion_indicator.ABSENT", "EMT_score","days_to_last_known_alive",
             "vital_status.Dead",
             "age_at_initial_pathologic_diagnosis")
Lasso_st = x_dummy[,selected]

# Multivariate cox model
#summary(coxph(Surv(days_to_last_known_alive, vital_status.Dead)~., data = Lasso_st))
summary(os <- coxph(Surv(days_to_last_known_alive, vital_status.Dead)~Hys_LN.LNneg+clinStage.II+lymphovascular_invasion_indicator.ABSENT+EMT_score+age_at_initial_pathologic_diagnosis, data = Lasso_st))
```

significance variable: Hys_LN.LNneg, clinStage.II, lymphovascular_invasion_indicator.ABSENT, EMT_score


### Survival analysis plot
```{r}
#library(survminer)
ggsurvplot(survfit(Surv(days_to_last_known_alive, vital_status.Dead)~Hys_LN.LNneg, data = Lasso_st)) 
ggsurvplot(survfit(Surv(days_to_last_known_alive, vital_status.Dead)~clinStage.II, data = Lasso_st))
ggsurvplot(survfit(Surv(days_to_last_known_alive, vital_status.Dead)~lymphovascular_invasion_indicator.ABSENT, data = Lasso_st))
#ggsurvplot(survfit(Surv(days_to_last_known_alive, vital_status.Dead)~EMT_score, data = Lasso_st))
#table(data_surv$clinStage, data_surv$vital_status)
```

### Test significance for cox model
```{r}
#summary(os <- coxph(Surv(days_to_last_known_alive, vital_status.Dead)~., data = Lasso_st))
cox.zph(os)
```

p>0.05, follow proportianal hazards, and vise versa.